Challenges and Solutions of Developing and Implementing a Novel Desktop-as-a-Service

### 2024-05-10 paper.tex
Minor spellings in chapter architecture

\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering,\\
	Frankfurt University of Applied Sciences,\\
	Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
	\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
	Desktop-as-a-Service (DaaS) is a cloud service category
	with tremendous potential for improving the everyday work
	of running and using applications.
	However, existing DaaS solutions have several drawbacks,
	which have limited the popularity of this service category in the past.
	Our novel DaaS solution, DESIGN, overcomes existing solutions limitations
	and offers significant benefits.
	It can potentially revolutionize how we use applications,
	supporting non-modified Linux and Windows applications
	with just a browser on the client side.
	This paper describes some of the knowledge we gained
	during the development because of the challenges
	we faced and the solutions we found.

	\keywords{DaaS \and Compatibility \and Performance \and Stability \and Usability}
\end{abstract}
%
%
%
\section{Introduction}

This paper discusses challenges during the development
and implementation of a novel Desktop-as-a-Service (DaaS) solution
called DESIGN that enables the deployment
and usage of unmodified Linux and Windows applications
in the same way as web applications
and can run inside public resources
but can also be deployed in a private context.
Since all interaction with DESIGN and the hosted applications
is done via a user's web browser,
the users can use any client, no matter what hardware or host operating system
it is compromised of.
A browser is the only software component required to use the novel DaaS service.

@@

\section{Related Work}
\label{sec:relatedworkAbschnitt}

@@

Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using the IaaS solution OpenStack~\cite{sefraoui2012openstack} and analyzed the characteristics and performance aspects of using noVNC, which is a client for the protocol Virtual Network Computing (VNC) that is implemented in the Hypertext Markup Language (HTML) and JavaScript, the protocol SPICE (Simple Protocol for Independent Computing Environments), and the remote desktop gateway Apache Guacamole as solutions for providing access to the desktop via a browser. One focus of the paper is the redirection of the sound interface of the virtual desktop. The paper's authors conclude that Guacamole, in combination with the Remote Desktop Protocol (RDP), is the best solution. The authors evaluated the lag time for remote audio playback in an Internet scenario by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via the monitoring software Wireshark. When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms. When using Guacamole combined with the protocol VNC, the average lag was around 1750\,ms. The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application. Again, the configuration Guacamole and RDP offered the best performance video update with an average lag of around 300\,ms. The configuration of Guacamole and VNC caused an average lag of around 500\,ms.

@@

\section{Architecture}
\label{sec:RecommendDaaSarchitecture}
% 
Figure~\ref{figure_architecture} shows the architecture of our novel DaaS solution DESIGN that can run unmodified Linux and Windows applications in Linux containers (Linux and Windows applications compatible with Wine) and virtual machines (Windows applications incompatible with Wine).
\begin{figure}
	\centering
	\includegraphics[width=.75\linewidth]{images/DaaS_DESIGN_Architecture_v12_english.pdf}
	\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.}
	\label{figure_architecture}
\end{figure}

@@

One of the main goals of DESIGN is to allow all interactions to be carried out solely using a browser. Thus, a broker or proxy software that mediates between the VNC or RDP implementation and the browser is required. The most advanced solution available is the free software project Apache Guacamole, which supports VNC and RDP and includes further relevant features regarding sound and printing.

@@

% TODO: Die letzten 3 Abschnitte noch überarbeiten und präzisieren

\section{Challenges and Solutions}
\label{sec:AnalysisPossibleComponents}

>>>>

\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering,\\
	Frankfurt University of Applied Sciences,\\
	Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
	\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
\maketitle              % typeset the header of the contribution

\begin{abstract}
	Desktop-as-a-Service (DaaS) is a cloud service category
	with tremendous potential for improving the everyday work
	of running and using applications.
	However, existing DaaS solutions have several drawbacks,
	which have limited the popularity of this service category in the past.
	Our novel DaaS solution, DESIGN, overcomes existing solutions limitations
	and offers significant benefits.
	It can potentially revolutionize how we use applications,
	supporting non-modified Linux and Windows applications
	with just a browser on the client side.
	This paper describes some of the knowledge we gained
	during the development because of the challenges
	we faced and the solutions we found.

	\keywords{DaaS \and Compatibility \and Performance \and Stability \and Usability}
\end{abstract}

\section{Introduction}
This paper discusses challenges during the development
and implementation of a novel Desktop-as-a-Service (DaaS) solution
called DESIGN that enables the deployment
and usage of unmodified Linux and Windows applications
in the same way as web applications
and can run inside public resources
but can also be deployed in a private context.
Since all interaction with DESIGN and the hosted applications
is done via a user's web browser,
the users can use any client, no matter what hardware or host operating system
it is compromised of.
A browser is the only software component required to use the novel DaaS service.

@@

\section{Related Work}\label{sec:relatedworkAbschnitt}

@@

Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS
using the IaaS solution OpenStack~\cite{sefraoui2012openstack}
and analyzed the characteristics and performance aspects of using noVNC,
which is a client for the protocol Virtual Network Computing (VNC)
that is implemented in the Hypertext Markup Language (HTML) and JavaScript,
the protocol SPICE (Simple Protocol for Independent Computing Environments),
and the remote desktop gateway Apache Guacamole as solutions
for providing access to the desktop via a browser.
One focus of the paper is the redirection
of the sound interface of the virtual desktop.
The paper's authors conclude that Guacamole,
in combination with the Remote Desktop Protocol (RDP), is the best solution.
The authors evaluated the lag time for remote audio playback
in an Internet scenario by measuring the time between the automated initiation
of the live streaming using the media player software VLC
and the arrival of the music data by analyzing network traffic
via the monitoring software Wireshark.
When using Guacamole combined with the protocol RDP,
the average lag was around 750\,ms.
When using Guacamole combined with the protocol VNC,
the average lag was around 1750\,ms.
The paper also includes a short evaluation of the lag for video frames
represented by background color changes of a Java application.
Again, Guacamole and RDP offered
the best video performance with an average lag of around 300\,ms.
The configuration of Guacamole and VNC caused an average lag of around 500\,ms.

@@

\section{Architecture}\label{sec:RecommendDaaSarchitecture}
% 
Figure~\ref{figure_architecture} shows the architecture
of our novel DaaS solution DESIGN
that can run unmodified Linux and Windows applications in Linux containers
(Native Linux applications and Native Windows applications compatible with Wine)
and virtual machines (Native Windows applications incompatible with Wine).
\begin{figure}
	\centering
	\includegraphics[width=.75\linewidth]{images/DaaS_DESIGN_Architecture_v12_english.pdf}
	\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.}
	\label{figure_architecture}
\end{figure}

@@

One of the main goals of DESIGN is to allow all interactions to be carried out
solely using a browser.
Thus, a broker or proxy software that mediates between the VNC or RDP
implementation and the browser is required.
The most advanced solution available
is the free software project Apache Guacamole, which supports VNC and RDP
and includes further relevant features regarding sound and printer support.

@@

\section{Challenges and Solutions}\label{sec:AnalysisPossibleComponents}
